---
title: "Exercise 2"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, simon.knuesel(at)wsl.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
```
<br>

### 1. Manipulate `airquality` dataset

i) Load the `airquality` dataset

ii) Rename the column headers to lower case

iii) Add a column with the variable `year`

iv) Create a new `date` column in the format (`YYYY-MM-DD`, e.g. `2019-09-25`)

v) *Advanced:* Try to do `i)-iv)` with `dplyr` and piping (`%>%`)
<br><br>


### 2. Convert table from `wide` to `long` format

i) Load the file `tree_growth_data_wide.rds` from the `01_Data` folder and 
give it a name (e.g. `wide_table`)

ii) Install the package `tidyr`

iii) Convert the table to the format shown below using the function 
`pivot_longer` from the `tidyr` package

```{r convert to long format, echo = FALSE, results = "hide"}
wide_table <- readRDS("/Users/simonknusel/Documents/02_Job/06_PostDoc/06_R-Courses/01_UZH_courses/2019/R_Basic_Introduction/01_Data/tree_growth_data_wide.rds")

long_table <- pivot_longer(data = wide_table, cols = 2:4, names_to = "series",
                           values_to = "value") %>% 
  arrange(series)
```
```{r head of long_table, echo = FALSE}
head(long_table)
```

iv) Save the table to the `01_Data` folder with a new name (e.g. `long_table`)


```{r pagebreak, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

## Hints

### 1. Manipulate `airquality` dataset

i) Use the command `data("airquality")` to load the dataset.

ii) Use the function `rename` (`?rename`) from the `dplyr` package.  
$\rightarrow$ Do not forget to load the `dplyr` package with `library(dplyr)`

iii) See `?airquality` to find out in which year the measurements were taken. 
Use the function `mutate` to add the new column `year`.

iv) Use the function `paste` (with `sep = "-"`) to combine the columns `year`, 
`month` and `day`.  
Use the function `as.Date` (with `format = "%Y-%m-%d"`) to create the new date 
column.

v) Build a structure similar to:  
```{r sample piping structure, echo = TRUE, eval = FALSE}
new_data_frame <- airquality %>%
  function() %>%  # add year column
  function() %>%  # combine year, month and day columns
  function()      # add date column
```
<br><br>


### 2. Convert table from `wide` to `long` format

i) Use the function `readRDS` (`?readRDS`).

ii) See the file `04_Install_packages.R` in the `03_Scripts` folder.

iii) Have a look at the vignette of the function with the command 
`vignette("pivot")`. A vignette is a help file with examples included.

iv) Use the function `saveRDS` to save the table as an `R` object to file.  
$\rightarrow$ Do not forget the file extension `.rds` at the end of the name


