---
title: "Exercise 1"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, simon.knuesel(at)wsl.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br>

### How to solve problems in R? - A suggestion...
Do I know what to do?

  + Yes:  
  $\rightarrow$ *Go for it...*
  + Maybe:  
  $\rightarrow$ *Try...*
  + No:  
  $\rightarrow$ *Look for help in R (`?function_name`, e.g. `?mean`)*  
  $\rightarrow$ *Ask Google (e.g. 'r how to calculate mean of data frame column')*  
  $\rightarrow$ *Look for help on one of the cheat sheets*  
  $\rightarrow$ *Look for hints on the last page of the exercise sheet*
<br><br>

### 1. Create vectors
Try to recreate the following vectors:

```{r create vec1, echo = FALSE, results = "hide"}
vec1 <- seq(from = 1, to = 10, by = 0.5)
```
```{r print vec1, echo = TRUE}
vec1
```
<br>
```{r create vec2, echo = FALSE, results = "hide"}
vec2 <- rep(c(1, 4, 8, 13), each = 4)
```
```{r print vec2, echo = TRUE}
vec2
```
<br><br>

### 2. Combine vectors
Combine the vectors `canton` and `peak` to `peak_canton`.

```{r create canton, echo = FALSE, results = "hide"}
canton <- c("GR", "TI", "UR", "BE", "VS")
```
```{r print canton, echo = TRUE}
canton
```

```{r create peak, echo = FALSE, results = "hide"}
peak <- c("Piz Bernina", "Adula Rheinwaldhorn", "Dammastock", 
          "Finsteraarhorn", "Dufourspitze")
```
```{r print peak, echo = TRUE}
peak
```
<br>
```{r create peak_canton, echo = FALSE, results = "hide"}
peak_canton <- paste(peak, canton, sep = "_")
```
```{r print peak_canton, echo = TRUE}
peak_canton
```
<br><br>

### 3. Load and save a .csv-file

i) Load the file `data_L1.csv` from the folder `01_Data` and give it a name (e.g. `my_table`)

ii) Save the object `my_table` as `my_table.csv` to the folder `01_Data`


```{r pagebreak, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

## Hints

### 1. Create vectors
Have a look at the functions `seq` and `rep`.  
You can see the help files by typing `?sep` and `?rep`.
<br>

### 2. Combine vectors
Have a look at the function `paste` (`?paste`).  
Check the argument `sep = " "` inside the function `paste`.
<br>

### 3. Load and save a .csv-file
Check the R-file `03_Load_and_save_data.R` in the `03_Scripts` folder or check the functions `read.csv` and `write.csv`.  

