---
title: "Solution to exercise 3"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, simon.knuesel(at)wsl.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br>

### 1. Plot the airquality dataset

i) Load the `airquality` dataset

```{r load airquality dataset, echo=TRUE, error=FALSE, warning=FALSE}
data("ariquality")
```

ii) Try to reproduce the plot shown below with `Temperature` on the x-axis
and `Ozone` on the y-axis

```{r plot airquality, echo = TRUE, results = "hide", fig.show = "hide"}
# code with comments
plot(data = airquality, Ozone ~ Temp, # define dataset and x and y variable
     las = 1, # horizontal axis numbers
     pch = 16, # change symbol of points
     xlab = "Temperature (°F)", ylab = "Ozone (ppb)", # change axis labels
     main = "Ozone vs. Temperature") # add a title
```
```{r plot airquality nocomments, echo = TRUE}
# code without comments
plot(data = airquality, Ozone ~ Temp, las = 1, pch = 16,
     xlab = "Temperature (°F)", ylab = "Ozone (ppb)",
     main = "Ozone vs. Temperature")
```

```{r pagebreak #1, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

iii) Calculate the median ozone concentration and add it to the plot as a line

```{r add mean wind speed, echo = TRUE}
# recreate plot
plot(data = airquality, Ozone ~ Temp, las = 1, pch = 16,
     xlab = "Temperature (°F)", ylab = "Ozone (ppb)",
     main = "Ozone vs. Temperature")

# calculate median ozone level
ozone_med <- median(airquality$Ozone, na.rm = TRUE)
# add median ozone level to plot
abline(h = ozone_med, col = "blue")
```

```{r pagebreak #2, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

iv) Add a second plot on the right with `Temperature` on the x-axis and
`Wind` on the y-axis

```{r two column plot, echo = TRUE, fig.height = 4, fig.width = 6}
# create plot window with two columns
par(mfrow = c(1, 2))

plot(data = airquality, Ozone ~ Temp, las = 1,
     pch = 16, xlab = "Temperature (°F)", ylab = "Ozone (ppb)",
     main = "Ozone vs. Temperature")

plot(data = airquality, Wind ~ Temp, las = 1,
     pch = 16, xlab = "Temperature (°F)", ylab = "Wind (mph)",
     main = "Wind vs. Temperature")
```

```{r pagebreak #3, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

v) Save these plots as a PDF

```{r save plots, echo = TRUE}
# 1) Define the name and size of the plot and where to save it
pdf(file = "airquality_plot.pdf", width = 8, height = 4)

# 2) Draw plots (i.e. run plot commands)
par(mfrow = c(1, 2))

plot(data = airquality, Ozone ~ Temp, las = 1,
     pch = 16, xlab = "Temperature (°F)", ylab = "Ozone (ppb)",
     main = "Ozone vs. Temperature")

plot(data = airquality, Wind ~ Temp, las = 1,
     pch = 16, xlab = "Temperature (°F)", ylab = "Wind (mph)",
     main = "Wind vs. Temperature")

# 3) Write plot to file by shutting off the plotting device
dev.off()
```

```{r pagebreak #4, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

vi) Calculate a linear regression between wind and temperature and add the 
regression line to the respective plot

```{r calculate lm, echo = TRUE, fig.width = 6, figh.height = 4}
# calculate linear model
lm <- lm(formula = Wind ~ Temp, data = airquality)
# inspect the summary of the linear model
summary(lm)

# recreate the plots
par(mfrow = c(1, 2))

plot(data = airquality, Ozone ~ Temp, las = 1,
     pch = 16, xlab = "Temperature (°F)", ylab = "Ozone (ppb)",
     main = "Ozone vs. Temperature")

plot(data = airquality, Wind ~ Temp, las = 1,
     pch = 16, xlab = "Temperature (°F)", ylab = "Wind (mph)",
     main = "Wind vs. Temperature")

# add the linear model to the plot
# (add 'abline' after the plot in which it should be displayed)
abline(lm, col = "red")
```

