---
title: "Solution to exercise 2"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, simon.knuesel(at)wsl.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
```

## Data exploration

### Explore the `airquality` dataset

i) Load the `airquality` dataset
```{r load airquality, echo = TRUE, results = "hide"}
data(airquality)
```
```{r head of airquality, echo = TRUE}
head(airquality)
```
<br>

ii) Assign a new name to the dataset (e.g. `airquality_1`)
```{r assign new name, echo = TRUE}
airquality_1 <- airquality
```
<br>

iii) Get an overview of the dataset:  
How many rows does the dataset have?
```{r number of rows, echo = TRUE}
dim(airquality_1) # returns the number of rows and columns

nrow(airquality_1) # returns the number or rows
```
<br>

How many columns does the dataset have?
```{r number of columns, echo = TRUE}
dim(airquality_1) # returns the number of rows and columns

ncol(airquality_1) # returns the number or columns
```


```{r pagebreak #1, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```


What class do the columns have? Can you guess?
```{r column classes, echo = TRUE}
class(airquality_1$Ozone)
class(airquality_1$Solar.R)
class(airquality_1$Wind)
```
<br>

iv) Calculate the mean temperature
```{r calculate mean temperature, echo = TRUE}
# select column Temp
temp <- airquality_1$Temp # select with name
temp <- airquality_1[, 4] # select with number

# calculate the mean
mean(temp)

# or combined in one line
mean(airquality_1$Temp)
```
<br>

v) What is the maximum ozone content?
```{r max ozone content, echo = TRUE}
# select column Ozone
ozone <- airquality_1$Ozone
ozone <- airquality_1[, 1]

# calculate the maximum
max(ozone) # result is NA because ozone contains NA's
max(ozone, na.rm = NA)

# or combined in one line
max(airquality_1$Ozone, na.rm = TRUE)
```


```{r pagebreak #2, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```


vi) In which month and day did the maximum ozone content occur?
```{r which.max ozone content, echo = TRUE}
# select column Ozone
ozone <- airquality_1$Ozone

# option 1
which.max(ozone) # row number of maximum
airquality_1[117, ] # select the row with the maximum ozone measurement

# option 2
airquality_1[which.max(ozone), ]

# option 3
# select the row with the maximum ozone measurement and the
# Month and Day column
airquality_1[117, 5:6]
```
<br>

vii) Calculate the mean temperature for the month May
```{r temp may, echo = TRUE}
# subset to the month May
airquality_may <- subset(x = airquality_1, Month == 5)
unique(airquality_may$Month) # check the result

# select the column Temp
temp_may <- airquality_may$Temp

# calculate the mean
mean(temp_may)
```
