---
title: "Solution to exercise 3"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, simon.knuesel(at)wsl.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data exploration

### Explore the `temperature` dataset

i) Load the dataset `temperature.csv` from the `01_Data` folder and assign
   it to an object with a meaningful name (e.g. `temperature`)
   
```{r set working directory, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/simonknusel/Documents/02_Job/06_PostDoc/06_R-Courses/01_UZH_courses/2019/R_Basic_Introduction/01_Data/")
```
```{r set working directory #2, eval = FALSE}
# Set the working directory
setwd("~/R_Basic_Introduction/01_Data") # replace with your path to the folder "01_Data"
```
```{r load temperature dataset}
# Load data
temperature <- read.csv(file = "temperature.csv")
```
```{r reset working directory, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/simonknusel/Documents/02_Job/06_PostDoc/06_R-Courses/01_UZH_courses/2019/R_Basic_Introduction/06_Solutions/")
```
<br>

ii) Get an overview of the dataset:  
View the first six rows of the dataset
```{r head of dataset}
head(temperature)
```

<br>

How many rows does the dataset have?
```{r number of rows}
nrow(temperature) # returns the number or rows

dim(temperature) # returns the number of rows and columns
```

<br>

How many columns does the dataset have?
```{r number of columns}
ncol(temperature) # returns the number or columns

dim(temperature) # returns the number of rows and columns
```


```{r pagebreak #1, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```


What class do the columns have? Can you guess?
```{r column classes}
class(temperature$site)
class(temperature$temp)
class(temperature$day)
class(temperature$month)
```

<br>

*All in one*: Sturcture of the dataset
```{r structure of dataset}
str(temperature)
```
<br>

iii) Calculate the mean temperature
```{r calculate mean temperature}
# select column 'temp'
temp <- temperature$temp # select with name
temp <- temperature[, 2] # select with number

# calculate the mean
mean(temp) # result is NA because 'temp' contains NA's
mean(temp, na.rm = TRUE)

# or combined in one line
mean(temperature$temp, na.rm = TRUE)
```


```{r pagebreak #2, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```


iv) In which months were the measurements taken?
```{r temp measurement months}
# extract unique values of column 'month'
month_measure <- unique(temperature$month)
month_measure
```
<br>

v) What month and day was the maximum temperature measured?
```{r which.max temperature}
# select column 'temp'
temp <- temperature$temp

# option 1
which.max(temp) # row number of maximum
temperature[180, ] # select the row with the maximum temperature measurement

# option 2
temperature[which.max(temp), ]

# option 3
# select the row with the maximum temperature measurement and the columns
# 'Month' and 'Day'
temperature[180, 3:4]
```


```{r pagebreak #3, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```


vi) Load the internal dataset `airquality` and caluclate the Pearson 
correlation between `Wind` and `Temp`. Do you expect a positive or negative
correlation?
```{r pearson correlation}
# load internal dataset airquality
data(airquality)

# get an overview of the dataset
head(airquality)
str(airquality)

# calculate the Pearson correlation
cor_wind_temp <- cor(airquality$Wind, airquality$Temp, method = "pearson")
cor_wind_temp
```
