---
title: "Solution to exercise 4"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, simon.knuesel(at)wsl.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data manipulation

### Manipulate the `temperature` dataset

i) Load the dataset `temperature.csv` from the `01_Data` folder and
assign it to an object with a meaningful name
```{r set working directory, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/simonknusel/Documents/02_Job/06_PostDoc/06_R-Courses/01_UZH_courses/2019/R_Basic_Introduction/01_Data/")
```
```{r set working directory #2, eval = FALSE}
# Set the working directory
setwd("~/R_Basic_Introduction/01_Data") # replace with your path to the folder "01_Data"
```
```{r load temperature dataset}
# Load data
temperature <- read.csv(file = "temperature.csv")

# Get an overview of dataset
head(temperature) # show the first six rows
tail(temperature) # show the last six rows
str(temperature) # show the structure of the dataset
```
```{r reset working directory, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/simonknusel/Documents/02_Job/06_PostDoc/06_R-Courses/01_UZH_courses/2019/R_Basic_Introduction/06_Solutions/")
```


```{r pagebreak #1, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```


ii) Filter the dataset for the site `Bern`
```{r filter site}
# Filter for site Bern
temp_be <- subset(temperature, site == "Bern")
unique(temp_be$site) # check the result
```
<br>

iii) Add a column with the variable `year` (the year is `2013`)
```{r add year column}
# Add year column
temp_be$year <- 2013
head(temp_be) # check the result
```
<br>

iv) Create a new `date` column
    - Create a new column that is a combination of the variables `year`, 
      `month` and `day` (in the form of `"2013-01-25"`)
    - Convert the `class` of the column from `"character"` to `"Date"`
    - Calculate the number of days between the first and last measurement
```{r create date column}
# Create a date vector
date_vec <- paste(temp_be$year, temp_be$month, temp_be$day, sep = "-")
head(date_vec)

# Add the date vector as a column to the dataset
temp_be$date <- date_vec
head(temp_be)
```


```{r pagebreak #2, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```


```{r create date column #2}
# Change the class of the 'date' column to "Date"
class(temp_be$date)
temp_be$date <- as.Date(temp_be$date)
class(temp_be$date)

# Calculate the number of days between first and last measurement
time_diff <- max(temp_be$date) - min(temp_be$date)
time_diff
```
<br><br>

v) Calculate the average temperature for periods without frost (i.e. 
the temperature is above `0 °C`) for the site `Zurich`
```{r average temperature}
# Filter the dataset for the site Zurich
temp_zh <- subset(temperature, site == "Zurich")

# Filter the dataset for periods without frost
temp_zh_nofrost <- subset(temp_zh, temp > 0)

# Example of a combination of the two filter arguments above
temp_zh_nofrost <- subset(temp_zh, site == "Zurich" & temp > 0)

# Calculate the average temperature
mean(temp_zh_nofrost$temp)
```


```{r pagebreak #3, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```


vi) Load the internal dataset `airquality` and change the column names to
lower case
```{r lower case column names}
# Load airquality dataset
data(airquality)

# Assign the dataset to a new object
airquality_1 <- airquality

# Option 1: manual change
colnames(airquality_1) <- c("ozone", "solar.r", "wind", "temp", "month", 
                            "day")
head(airquality_1) # check the result
```

To find a function for a specific task it is usually helpful to Google for
answers. For this specific exercise I would use the following query:  
*r convert column names to lower case*

```{r lower case column names #2}
# Option 2: use function 'tolower'
?tolower
colnames(airquality_1) <- tolower(colnames(airquality))
head(airquality_1) # check the result
````
