---
title: "Solution to Exercise 1"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, simon.knuesel(at)wsl.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What results do you expect of the following commands?
```{r guess results solution #1}
x <- c(2, 5, 6, 5)
y <- c(3, 5, 8)

?class
class(x)

x + 1

x + y
```

Explanation to `x + y`:  
When you add vectors, the first element of the first vector is added to the
first element of the second vector 
(i.e. `x + y = c(x[1] + y[1], x[2] + y[2], etc.)`).
If one vector is longer than the other (in this case `x`), the shorter
vector is recycled (i.e. `(x + y)[4] = x[4] + y[1]`).

```{r guess results solution #2}
y[2:3]
x[x > 5]

x <- x[1:2]
length(x)

member <- c(TRUE, TRUE, FALSE, TRUE)
?sum
sum(member)

shoe_size <- c(35, 42, 44, 36, 38, 39)
length(shoe_size)

mean(shoe_size)

median(shoe_size)

?which.max
which.max(shoe_size)
```

Explanation to `which.max()`:  
The function `which.max()` returns the position of the element with the maximum
value inside a vector. In this case, the maximum value is `44`, which is at the 
third position inside the vector `shoe_size`. Therefore, 
`which.max(shoe_size)` returns the value `3`.
