---
title: "Exercise 5"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, r\_courses(at)gmx.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data visualisation

### Visualise the `airquality` dataset

i) Load the internal dataset `airquality`

ii) Try to reproduce the plot shown below with `Temperature` on the x-axis
and `Ozone` on the y-axis

```{r plot airquality, echo = FALSE, fig.height = 3, fig.width = 5}
data("airquality")

plot(data = airquality, Ozone ~ Temp, las = 1,
     pch = 16, xlab = "Temperature (°F)", ylab = "Ozone (ppb)",
     main = "Ozone vs. Temperature")
```
<br>

iii) Calculate the median ozone concentration and add it to the plot as a line

```{r add mean wind speed, echo = FALSE, fig.height = 3, fig.width = 5}
plot(data = airquality, Ozone ~ Temp, las = 1,
     pch = 16, xlab = "Temperature (°F)", ylab = "Ozone (ppb)",
     main = "Ozone vs. Temperature")

# calculate median ozone level
ozone_med <- median(airquality$Ozone, na.rm = TRUE)
abline(h = ozone_med, col = "blue")
```

```{r pagebreak #1, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

iv) Add a second plot on the right with `Temperature` on the x-axis and
`Wind` on the y-axis

```{r two column plot, echo = FALSE, fig.height = 4, fig.width = 6}
par(mfrow = c(1, 2))

plot(data = airquality, Ozone ~ Temp, las = 1,
     pch = 16, xlab = "Temperature (°F)", ylab = "Ozone (ppb)",
     main = "Ozone vs. Temperature")

plot(data = airquality, Wind ~ Temp, las = 1,
     pch = 16, xlab = "Temperature (°F)", ylab = "Wind (mph)",
     main = "Wind vs. Temperature")
```
<br>

v) Save these plots as a PDF
<br><br>

### *For those who have time left:*

vi) Calculate a linear regression between wind and temperature and add the 
regression line to the respective plot


```{r pagebreak #2, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

## Hints

i) Load the `airquality` dataset
   - Use the command `data(airquality)` to load the dataset.

<br>

ii) Try to reproduce the plot shown below with `Temperature` on the x-axis
and `Ozone` on the y-axis
    - Axis labels: Check the arguments `xlab` and `ylab` (`?xlab, ?ylab`)
    - Title: Check the argument `main` (`?main`)
    - Horizontal axis numbering: Check the argument `las` (`?las`)

<br>

iii) Calculate the median ozone concentration and add it to the plot as a line
     - Use the command `median` to calculate the median  
       $\rightarrow$ Do not forget to remove `NA` values with `na.rm = TRUE`
     - Use the function `abline` (`?abline`) to add a line to an existing plot

<br>

iv) Add a second plot on the right with `Temperature` on the x-axis and
`Wind` on the y-axis
    - Use the command `par(mfrow = c(1, 2))` to create two plot windows (`c(1, 2)`
      means 1 row and 2 columns).

<br>

v) Save these plots as a PDF  
    Use the following structure:
    1. `pdf(file = "path_to_file/file_name.pdf")`
    2. Code to produce the plots (as many lines of code as needed)
    3. `dev.off()`

<br>

vi) Calculate a linear regression between wind and temperature and add the 
regression line to the respective plot
    - Use the function `lm` (`?lm`) to calculate a linear regression
    - Use the function `abline` (`?abline`) to add the regression line to the plot

