---
title: "Exercise 3"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, simon.knuesel(at)wsl.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
```

## Data manipulation

### Manipulate the `temperature` dataset

i) Load the dataset `temperature.csv` from the `01_Data` folder and
give it a new name (e.g. `temp`)

<br>

ii) Filter the dataset for the site `Bern`

<br>

ii) Add a column with the variable `year` (the year is `2013`)

<br>

*For those who have time left...*

iv) Create a new `date` column
    - Create a new column with the variables `year`, `month` and `day` combined
      (in the form of `"2013-01-25"`)
    - Convert the `class` of the column from `"character"` to `"date"`
  
<br>

v) Calculate the average temperature for periods without frost (i.e. 
the temperature is above `0 °C`) for the site `Zurich`


```{r pagebreak, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

## Hints

i) Load the dataset `temperature.csv` from the `01_Data` folder and
give it a new name
   - Set the working directory to the `01_Data` folder in the course material
     (`?setwd` or look at the slides in the `02_Slides` folder)
   - Load the dataset with the function `read.csv` (`?read.csv`)

<br>

ii) Check the function `subset` (`?subset`)

<br>

ii) Use a command in the form of `dataframe$year` and assign the value `2013`
to this column.

<br>

*For those who have time left...*

iv) Create a new `date` column
    - Create a vector that combines the columns `year`, `month` and `day`.
      Check the function `paste` (`?paste`) and the argument `sep` inside the
      function `paste`
    - Add this vector to the dataset and name the column `date`
    - Check the function `as.Date` (`?as.Date`) to change the class of the
      column from `"character"` to `"Date"`
  
<br>

v) Calculate the average temperature for periods without frost (i.e. 
the temperature is above `0 °C`) for the site `Zurich`
   - Filter the dataset with the function `subset` (`?subset`)
   - Calculate the average temperature with the function `mean` (`?mean`)

