---
title: "Exercise 3"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, simon.knuesel(at)wsl.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data exploration

### Explore the `temperature` dataset

i) Load the dataset `temperature.csv` from the `01_Data` folder and assign
   it to an object with a meaningful name (e.g. `temperature`)
<br>

ii) Get an overview of the dataset:
    - View the first six rows of the dataset
    - How many rows does the dataset have?
    - How many columns does the dataset have?
    - What class do the columns have? Can you guess?
<br><br>

iii) Calculate the mean temperature

<br>

### *For those who have time left...*

iv) In which months were the measurements taken?
<br>

v) What month and day was the maximum temperature measured?
<br>

vi) Load the internal dataset `airquality` and caluclate the Pearson 
correlation between `Wind` and `Temp`. Do you expect a positive or negative
correlation?


```{r pagebreak, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

## Hints

i) Load the dataset `temperature.csv` from the `01_Data` folder and assign
   it to an object with a meaningful name (e.g. `temp`)
   - Set the working directory to the `01_Data` folder in the course material
     (`?setwd` or look at the slides in the `02_Slides` folder)
   - Load the dataset with the function `read.csv` (`?read.csv`)

<br>

ii) Get an overview of the dataset:
    - First six rows: Check the function `head` (`?head`)
    - Number of rows: Check the functions `nrow` (`?nrow`) or `dim` (`?dim`)
    - Number of columns: Check the functions `ncol` (`?ncol`) or `dim` (`?dim`)
    - Column classes: Check the function `class` (`?class`)
    - *All in one*: Check the function `str` (`?str`)

<br>

iii) Check the function `mean` (`?mean`).  
     Select the column `temp` (in the form of `df$temp` or `df[, 2]`)

<br>

iv) Check the function `unique` (`?unique`)

<br>

v) Check the function `which.max` (`?which.max`)

<br>

vi) Load the dataset with `data(airquality)`  
    Check the function `cor` (`?cor`) and the argument `method`
