---
title: "Exercise 2"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, simon.knuesel(at)wsl.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
```

## Data exploration

### Explore the `airquality` dataset

i) Load the internal dataset `airquality`

<br>

ii) Assign a new name to the dataset (e.g. `airquality_1`)

<br>

iii) Get an overview of the dataset:
     - How many rows does the dataset have?
     - How many columns does the dataset have?
     - What class do the columns have? Can you guess?

<br><br>

iv) Calculate the mean temperature

<br>

v) What is the maximum measured ozone content?  

<br>

*For those who have time left...*

vi) In which month and day was the maximum ozone content measured?  

<br>

vii) Calculate the mean temperature for the month May


```{r pagebreak, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

## Hints

i) Use the command `data(airquality)` to load the dataset.

<br>

ii) Assign a new name with the form `new_name <- airquality`

<br>

iii) Get an overview of the dataset:
     - Number of rows: Check the functions `dim` (`?dim`) or `nrow` (`?nrow`)
     - Number of columns: Check the functions `dim` (`?dim`) or `ncol` (`?ncol`)
     - Column classes: Check the function `class` (`?class`)

<br>

iv) Check the function `mean` (`?mean`). Select the column `Temp` 
(in the form of `df$Temp` or `df[, 4]`)

<br>

v) Check the function `max` (`?max`) and the argument `na.rm`

<br>

vi) Check the function `which.max`

<br>

vii) Use the function `subset` (`?subset`) to filter for the month May 
(`Month == 5`)
