---
title: "Exercise 4"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, simon.knuesel(at)wsl.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data visualisation

### Plot the airquality dataset

i) Load the `airquality` dataset

ii) Try to reproduce the plot shown below with `Temperature` on the x-axis
and `Ozone` on the y-axis

```{r plot airquality, echo = FALSE, fig.height = 3, fig.width = 5}
data("airquality")

plot(data = airquality, Ozone ~ Temp, las = 1,
     pch = 16, xlab = "Temperature (°F)", ylab = "Ozone (ppb)",
     main = "Ozone vs. Temperature")
```
<br>

iii) Calculate the median ozone concentration and add it to the plot as a line

```{r add mean wind speed, echo = FALSE, fig.height = 3, fig.width = 5}
plot(data = airquality, Ozone ~ Temp, las = 1,
     pch = 16, xlab = "Temperature (°F)", ylab = "Ozone (ppb)",
     main = "Ozone vs. Temperature")

# calculate median ozone level
ozone_med <- median(airquality$Ozone, na.rm = TRUE)
abline(h = ozone_med, col = "blue")
```

```{r pagebreak #1, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

iv) Add a second plot on the right with `Temperature` on the x-axis and
`Wind` on the y-axis

```{r two column plot, echo = FALSE, fig.height = 4, fig.width = 6}
par(mfrow = c(1, 2))

plot(data = airquality, Ozone ~ Temp, las = 1,
     pch = 16, xlab = "Temperature (°F)", ylab = "Ozone (ppb)",
     main = "Ozone vs. Temperature")

plot(data = airquality, Wind ~ Temp, las = 1,
     pch = 16, xlab = "Temperature (°F)", ylab = "Wind (mph)",
     main = "Wind vs. Temperature")
```
<br>

v) Save these plots as a PDF
<br>

*For those who have time left:*

vi) Calculate a linear regression between wind and temperature and add the 
regression line to the respective plot


```{r pagebreak #2, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

## Hints

i) Load the `airquality` dataset
Use the command `data("airquality")` to load the dataset.

<br>

ii) Try to reproduce the plot shown below with `Temperature` on the x-axis
and `Ozone` on the y-axis
    - Axis labels: `xlab = "some text"`, `ylab = "some text"`
    - Title: `main = "some text"`
    - Horizontal axis numbering: `las = 1`

<br>

iii) Calculate the median ozone concentration and add it to the plot as a line
     - Use the command `median` to calculate the median ($\rightarrow$ Do not 
       forget to remove `NA` values with `na.rm = TRUE`)
     - Use the command `abline` to add a line to an existing plot

<br>

iv) Add a second plot on the right with `Temperature` on the x-axis and
`Wind` on the y-axis
Use the command `par(mfrow = c(1, 2))` to create two plot windows (`c(1, 2)`
means 1 row and 2 columns).  
Alternatively you can use the command 
`layout(mat = matrix(c(1:2), ncol = 2))`.

<br>

v) Save these plots as a PDF  
    Use the following structure:
    1. `pdf(file = "path_to_file/file_name.pdf")`
    2. Code to produce the plots (as many lines of code as needed)
    3. `dev.off()`

<br>

vi) Calculate a linear regression between wind and temperature and add the 
regression line to the respective plot
    - Use the command `lm(...)` (`?lm`) to calculate a linear regression
    - Use the command `abline` to add the regression line to the plot

