---
title: "Exercise 1"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, r_courses(at)gmx.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br>

I recommend you to watch the video tutorial 'Getting_started_with_RStudio'
in the folder '02_Slides_tutorials' of the course material before starting 
the exercise.
<br>


## What results do you expect of the following commands?

```{r guess results, results = "hide", warning = FALSE, error=FALSE}
x <- c(2, 5, 6, 5)
y <- c(3, 5, 8)

?class
class(x)

x + 1
x + y

y[2:3]
x[x > 5]

x <- x[1:2]
length(x)

member <- c(TRUE, TRUE, FALSE, TRUE)
?sum
sum(member)
```
<br><br>


### *For those who have time left...*

```{r guess results advanced, results = "hide", warning=FALSE, message=FALSE}
shoe_size <- c(35, 42, 44, 36, 38, 39)
length(shoe_size)

mean(shoe_size)

median(shoe_size)

which.max(shoe_size)
```


```{r pagebreak #1, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```


## Hints

### What results do you expect of the following commands?
Type the code into your console and execute the code.  
Shortcut Mac: Cmd + Enter  
Shortcut Windows: Ctrl + Enter
