---
title: "Exercise 4"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, simon.knuesel(at)wsl.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data manipulation

### Manipulate the `temperature` dataset

i) Load the dataset `temperature.csv` from the `01_Data` folder and
assign it to an object with a meaningful name

<br>

ii) Filter the dataset for the site `Bern`

<br>

iii) Add a column with the variable `year` (the year is `2013`)

<br>

### *For those who have time left...*

iv) Create a new `date` column
    - Create a new column that is a combination of the variables `year`, 
      `month` and `day` (in the form of `"2013-01-25"`)
    - Convert the `class` of the column from `"character"` to `"Date"`
    - Calculate the number of days between the first and last measurement
  
<br>

v) Calculate the average temperature for periods without frost (i.e. 
the temperature is above `0 °C`) for the site `Zurich`

<br>

vi) Load the internal dataset `airquality` and change the column names to
lower case


```{r pagebreak, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

## Hints

i) Load the dataset `temperature.csv` from the `01_Data` folder and
assign it to an object with a meaningful name
   - Set the working directory to the `01_Data` folder in the course material
     (`?setwd` or look at the video tutorial in the `02_Slides_tutorials` folder)
   - Load the dataset with the function `read.csv` (`?read.csv`)

<br>

ii) Check the function `subset` (`?subset`)

<br>

iii) Use a command in the form of `df$year` and assign the value `2013`
to this column.

<br>

iv) Create a new `date` column
    - Combine the columns `year`, `month` and `day`:  
      Check the function `paste` (`?paste`) and the argument `sep` inside the
      function `paste`
    - Add this vector to the dataset and name the column `date`
    - Check the function `as.Date` (`?as.Date`) to change the class of the
      column from `"character"` to `"Date"`
    - Calculate the duration of the measurements:  
      Calculations are possible with variables of class `"Date"`
  
<br>

v) Calculate the average temperature for periods without frost (i.e. 
the temperature is above `0 °C`) for the site `Zurich`
   - Filter the dataset with the function `subset` (`?subset`)
   - Calculate the average temperature with the function `mean` (`?mean`)

vi) Change the column names to lower case
    - Use the function `colnames` (`?colnames`) to change column names  
      $\rightarrow$ See file `02_Vector_data_frame_list.R` in the `03_Scripts`
      folder for an example
    - Try to do it manually and also try to find a function that does this task
      (e.g. by asking Google)

