---
title: "Solution to exercise 2"
header-includes: |
  \usepackage{fancyhdr}
  \usepackage{datetime}
  \newdateformat{mydate}{\the\day.\the\month.\the\year}
  \pagestyle{fancy}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[L]{\empty}
  \fancyhead[R]{\thepage}
  \fancyfoot[C]{
    \begin{tabular}[b]{c}%
      \small R: Basic Introduction, University of Zurich | \mydate\today\ \\
      \small by Simon Knüsel, simon.knuesel(at)wsl.ch \\
    \end{tabular}
  }
output:
  pdf_document: default
  html_document:
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
```
<br>

### 1. Manipulate `airquality` dataset

i) Load the `airquality` dataset
```{r load airquality, echo = TRUE, results = "hide"}
data(airquality)
```
```{r head of airquality, echo = TRUE}
head(airquality)
```
<br>

ii) Rename the column headers to lower case
```{r rename column headers #1, echo = TRUE}
# convert column names to lower case in the classical way
airquality1 <- airquality
colnames(airquality1) <- tolower(colnames(airquality1))

# convert column names to lower case with dplyr
library(dplyr)
# option 1
airquality1 <- rename(airquality, ozone = Ozone, solar.rad = Solar.R, 
                      wind = Wind, temp = Temp, month = Month, day = Day)
```
```{r rename column headers #2, echo = TRUE, results = "hide"}
# option 2 (advanced)
airquality1 < airquality %>% 
  # convert to lower case
  rename_all(tolower)

```
```{r rename column headers #3, echo = TRUE}
# Additional example:
# convert dot (.) to underscore (_)
# option 1
colnames(airquality1)[2] <- "solar_r"
# option 2 (advanced)
colnames(airquality1) <- gsub("\\.", "_", colnames(airquality1))

# Additional example with dplyr (advanced):
airquality1 <- airquality %>% 
  # convert to lower case
  rename_all(tolower) %>% 
  # replace dots (.) with underscores (_)
  rename_all(~gsub("\\.", "_", .))
```

```{r pagebreak #1, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

iii) Add a column with the variable `year`
```{r add year column, echo = TRUE}
# Look at the help file of the dataset airquality
?airquality

airquality_year <- mutate(airquality1, year = 1973)
head(airquality_year)
```
<br>


iv) Create a new `date` column in the format (`YYYY-MM-DD`, e.g. `2019-09-25`)
```{r create date column, echo = TRUE}
# Create the new column as vector
date_vec <- paste(airquality_year$year, airquality_year$month, 
                  airquality_year$day, sep = "-")
date_vec[1:6]

# Add the new column to airquality
airquality_date <- mutate(airquality1, date = date_vec)
class(airquality_date$date)

# Convert the date from class character to class date
airquality_date <- mutate(airquality_date, date = as.Date(date, format = "%Y-%m-%d"))
class(airquality_date$date)
str(airquality_date)
```

```{r pagebreak #2, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```

v) Try to do `i)-iv)` with `dplyr` and piping (`%>%`)
```{r dplyr piping, echo = TRUE}
# load original airquality dataset
data(airquality)

airquality2 <- airquality %>% 
  # rename columns
  rename(ozone = Ozone, solar_rad = Solar.R, wind = Wind, temp = Temp, 
         month = Month, day = Day) %>% 
  # add year column
  mutate(year = 1973) %>% 
  # add date column
  mutate(date = paste(year, month, day, sep = "-")) %>% 
  # change class of date column to date
  mutate(date = as.Date(date))

head(airquality2)
str(airquality2)
```

```{r pagebreak #3, results="asis", echo = FALSE}
if ("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
  cat("\\pagebreak")
}
```


### 2. Convert table from `wide` to `long` format

```{r set working directory, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/simonknusel/Documents/02_Job/06_PostDoc/06_R-Courses/01_UZH_courses/2019/R_Basic_Introduction/")
```

i) Load the file `tree_growth_data_wide.rds` from the `01_Data` folder and 
give it a name (e.g. `wide_table`)

```{r load tree_growth_data_wide, echo = TRUE}
# maybe you need to change the working directory or the file path
# (remember to include the filename extension '.rds')
wide_table <- readRDS("01_Data/tree_growth_data_wide.rds")
head(wide_table)
```
<br>

ii) Install the package `tidyr`
```{r install pck tidyr, eval = FALSE}
install.packages("tidyr")
```
```{r load tidyr library, echo = TRUE, message = FALSE, error = FALSE}
library(tidyr) # library needs to be loaded after the installation to be available
```
<br>

iii) Convert the table to the format shown below using the function 
`pivot_longer` from the `tidyr` package

```{r convert to long format #1, echo = TRUE}
long_table <- pivot_longer(data = wide_table, cols = 2:4, names_to = "series",
                           values_to = "value") %>% 
  # sort by series
  arrange(series)

head(long_table) # first six rows of a table

tail(long_table) # last six rows of a table
```

<br>

iv) Save the table to the `01_Data` folder with a new name (e.g. `long_table`)
```{r save long_table}
saveRDS(object = long_table, file = "01_Data/long_table.rds")
```

